### 2019-05-28 paper.tex
asdf

\documentclass[12pt,preprint,times]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{flafter}
%\usepackage{floatflt}
\usepackage{placeins}
\usepackage{siunitx}
\usepackage{graphicx}
\graphicspath{{./figures/}}% Include figure files
%\usepackage{epstopdf}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{appendix}
\usepackage{amsmath}
\usepackage{cases}
\usepackage{calc}
\usepackage{amssymb}
\usepackage[dvips]{color}
\usepackage{color}
\usepackage{enumitem}
\usepackage{soul}
\usepackage{indentfirst}
\usepackage{hyphenat}
\usepackage{xspace}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{lineno}
\usepackage{setspace}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
%\newcommand{\cmpersec}{$\frac{\text{cm}}{\text{s}}$~}
\newcommand{\cmpersec}{cm/s~}
\newcommand{\etal}{\textit{et al.}}
\newcommand{\na}{Na\textsuperscript{+}~}
\newcommand{\cl}{Cl\textsuperscript{-}~}
\newcommand{\PM}{$\pm$~}
\newcommand{\invnm}{$\text{\si{\nm}}^{-1}$~}
\newcommand{\nmeter}{\si{\nm}~}
\renewcommand{\arraystretch}{1.8}
\newcommand{\TODO}{\hl{TODO}~}
\newcommand{\about}{$\sim$}
\newcommand{\aangstroms}{{\aa}ngstr{\"o}ms}
\newcommand{\aangstrom}{{\aa}ngstr{\"o}m}
\newenvironment{hle}{\color{red}}

\newcommand{\beginsupplemental}{%
    \setcounter{table}{0}
    \renewcommand{\thetable}{S\arabic{table}}%
    \setcounter{figure}{0}
    \renewcommand{\thefigure}{S\arabic{figure}}%
}

\begin{document}
\bibliographystyle{plain}
\title{Need a title!!!! Interaction of salt with ether- and ester-linked phospholipid bilayers}
\author{Matthew Saunders} 
\maketitle
\tableofcontents
\listoffigures
\end{document}